Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-27 WSSSPE3/report/wg_appendix_publishing_SW.tex
Appendix K. No content change.
Appendix K:  A handful of changes.

A tentative first cut at the list containing executable papers identified the following:
\begin{itemize}
\item ACM Transactions on Mathematical Software (TOMS): provides the extra step
 of having reviewers validate the code which was submitted with the publication.
\item The Mathematica Journal: publishes Mathematica notebooks (with equations,
figures, etc.) directly.
\item O'Reilly Media: announced that it plans to make IPython Notebooks a first-class
 authoring environment for their publishing program alongside their existing mechanisms.
\item Nature: offering a list of notebooks published alongside more traditional articles,
 and is looking at ways to make these documents more official. There are, in fact, a
 number of journals that offer ``electronic supplements'' to the more traditionally published
 static articles.
\item IPython: maintained list of ``reproducible academic publications''~\cite{ipython-pubs}.
\item KBase: narratives build on IPython\slash Jupyter notebooks to build publications that are
  reproducible, and can be commented or annotated.
\end{itemize}

The group also discussed future possibilities for this new publication format which might
provide advantages:
\begin{itemize}
\item Journals could be built around an existing, widely used framework thereby reducing
  the burden of studying code on the part of reviewers (common bits of infrastructure
  which aren't relevant to the science would be automatically excluded).
\item Journals might be encouraged to use more metadata, making them easier to mine
  for various analytical purposes.
\item RIO is an effort to publish fragmentary results that can subsequently be combined
  into a single content item.
\item Papers could be made more understandable. Each equation or technical term could
  offer a link to a document/tutorial explaining its origin and/or derivation.
\item So many options for publication currently exist that good science may be getting
  lost in the noise. Would some sort of ``upvote'' mechanism be of value?
\item Some sort of Replicated Computation Results badge could be made available to
  publications that have undergone greater scrutiny (this is already done by TOMS).
\end{itemize}

@@

The opportunity is to collect a list of current executable papers and
shine a light on the experiments and development efforts currently underway.

The only obstacle to this is the difficulty in finding and identifying such
publications. The Software Sustainability Institute was able to do something similar
for publications about software by making a public page containing a catalog
of these publications and enlisting the help of the community to grow the list.

@@

Steven R. Brandt will create a first version of the page within a week or so of the WSSSPE3 conference.

>>>>

A tentative first cut at the list containing executable papers identified the following:
\begin{itemize}

\item ACM Transactions on Mathematical Software (TOMS): provides the extra step
of having reviewers validate the code which was submitted with the publication.

\item The Mathematica Journal: publishes Mathematica notebooks (with equations,
figures, etc.) directly.

\item O'Reilly Media: announced that it plans to make IPython Notebooks a
first-class authoring environment for their publishing program alongside their
existing mechanisms.

\item Nature: offering a list of notebooks published alongside more traditional
articles, and is looking at ways to make these documents more official. There
are, in fact, a number of journals that offer ``electronic supplements'' to the
more traditionally published static articles.

\item IPython: maintained a list of ``reproducible academic
publications''~\cite{ipython-pubs}.

\item KBase: narratives built on IPython\slash Jupyter notebooks; to build
publications that are reproducible, and can be commented or annotated.

\end{itemize}

The group also discussed future possibilities for a new publication format
which might provide advantages:
\begin{itemize}

\item Journals could be built around an existing, widely used framework thereby
reducing the burden of studying code on the part of reviewers (common bits of
infrastructure which are not relevant to the science would be automatically
excluded).

\item Journals might be encouraged to use more metadata, making them easier to
mine for various analytical purposes.

\item RIO is an effort to publish fragmentary results that can subsequently be
combined into a single content item. \choinote{What's RIO?}

\item Papers could be made more understandable. Each equation or technical term
could be linked to a document/tutorial explaining its origin and/or
derivation.

\item So many options for publication currently exist that good science may be
getting lost in the noise. Would some sort of ``upvote'' mechanism be of value?

\item Some sort of Replicated Computation Results badge could be made available
to publications that have undergone greater scrutiny (this is already done by
TOMS).

\end{itemize}

@@

The opportunity is to collect a list of executable papers and shine a
light on the experiments and development efforts currently underway.

The only obstacle to this is the difficulty in finding and identifying such
publications. The Software Sustainability Institute was able to do something
similar for publications about software by making a public page \choinote{where?} containing a
catalog of these publications and enlisting the help of the community to grow
the list.

@@

Steven R.\ Brandt will create a first version of the page within a week or so of
the WSSSPE3 conference.
